---
title: "iteration_and_listcols"
author: "Angelica Bailey"
date: "2025-10-28"
output: html_document
---


```{r}
library(tidyverse)
library(readxl)
library(rvest)
```


## Make a list
```{r}
l =
  list(
    vec_numeric = 1:23,
    char_vec = c("Jeff"),
    mat = matrix(1:8, nrow = 2, ncol = 4),
    summary = summary(rnorm(1000, mean = 4))
  )

l[[1]] #gives first element in list
l[["vec_numeric"]]
l$mat
```


## Make a different list
```{r}
list_normals =
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = 3, sd = 4)
  )
```


copy and paste the function from last time
```{r}
mean_and_sd = function(x) {
  
    if(!is.numeric(x)) {
    stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)
  
  #format output as table
  tibble(
    mean = mean_x,
    sd = sd_x
    )  
}
```

```{r}
mean_and_sd(list_normals[[1]])
mean_and_sd(list_normals[[2]])
mean_and_sd(list_normals[[3]])
mean_and_sd(list_normals[[4]])
```


Use a loop to iterate!

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_normals[[i]])
  
}
```


Use `map` to do the same thing

```{r}
output = map(list_normals, mean_and_sd)

output = map(list_normals, median)
```


check out `map` variants
```{r}
#creates dataframe
map_dfr(list_normals, mean_and_sd, .id = "sample")

#outputs single number to make collection of single numbers
map_dbl(list_normals, median)


```


## LIST COLUMNS

try to put my list into a datafram
```{r}
listcol_df =
  tibble(
    name = c("a","b","c","d"),
    sample = list_normals
  )
```


Did this really work?
```{r}
pull(listcol_df, sample)
```
Can I apply the mean and sd?
```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```

Iterate using `map`
```{r}
map(pull(listcol_df, sample), mean_and_sd)
```

Adding a column 
```{r}
listcol_df =
  listcol_df |> 
  mutate(
    summary = map(sample, mean_and_sd)
  )
```



Unnesting to show only mean and sd
```{r}
listcol_df |> 
  select(-sample) |> 
  unnest(summary)
```

## Revisit NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)


nsduh_import = function(html, table_num) {
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  data
}

nsduh_import(nsduh_html, table_num = 1)
```


Try this with for loop
```{r}
output = vector("list", length = 3)

for(i in 1:3) {
  
  output[[i]] = nsduh_import(html = nsduh_html, i)
}
```



Do this with `map`

```{r}
#two arguments that you have to specify
map(1:3, nsduh_import, html = nsduh_html)
```


Do this all in a dataframe

```{r}
nsduh_df =
  tibble(
    name = c("marj year", "marj month", "marj first"),
    number = 1:3
  ) |> 
  mutate(
    table = map(number, nsduh_import, html = nsduh_html)
  ) |> 
  unnest(table)
```


## Look at weather data

```{r}
library(p8105.datasets)
data("weather_df")
```

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() +
  geom_smooth(method = "lm")
```


Lets do a regression
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  lm(tmax ~ tmin, data = _)

weather_df |> 
  filter(name == "Molokai_HI") |> 
  lm(tmax ~ tmin, data = _)

weather_df |> 
  filter(name == "Waterhole_WA") |> 
  lm(tmax ~ tmin, data = _)
```


Lets iterate differently
```{r}
weather_nest =
  weather_df |> 
  nest(data = date:tmin)
```

```{r}
lm(tmax ~ tmin, data = pull(weather_nest, data)[[1]])
lm(tmax ~ tmin, data = pull(weather_nest, data)[[2]])
lm(tmax ~ tmin, data = pull(weather_nest, data)[[3]])
```

Using `map`
```{r}
weather_lm = function(df) {
  
  lm(tmax ~ tmin, data = df)
  
}



map(pull(weather_nest, data), weather_lm)
```


```{r}
weather_nest |> 
  mutate(
    lm_fits = map(data, weather_lm)
  )
```



